---
title: "Cluster Analysis"
author: "Daniel Ebbert"
date: "`r format(Sys.Date(), '%d.%m.%Y')`"
output: html_document
bibliography: Method.bib
---

```{r setup, include = FALSE}
# Set the knitr options
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      fig.path = "fig/")

# Load the required libraries
library(cluster)
library(dplyr)
library(labelled)
library(knitr)
library(tibble)
library(stringr)
library(haven)
library(anchors)
library(ggplot2)
library(scales)
library(reshape2)
library(forcats)
library(factoextra)
library(fpc)
library(conover.test)
library(chisq.posthoc.test)

# Source self written functions for this project as well as of the data submodule
source(paste(getwd(), "/R/00_functions.R", sep = ""))
source(paste(getwd(), "/data/R/functions.R", sep = ""))

# Set seed for reproducability
set.seed(5446)
```

## Load the data

First the original data is loaded, merged and cleaned.

```{r load_data}
# Merge and clean the datasets
source(paste(getwd(), "/R/01_merge-datasets.R", sep = ""))
```

## Average age and gender distribution

The average age in the full data set is: `r round(overall_average_age, digits =2)`
The standard deviation in the full data set is: `r round(overall_sd_age, digits =2)`
The gender distribution in the full data set is as follows:
```{r gender_distribution}
kable(gender_distribution)
```

## Distance matrix

Because the variables that are supposed to be part of the cluster analysis are not all continuous a different distance measure is needed. The used distance measure for this case is Gower distance [@Gower1971]. In order to be able to run a cluster analysis with a different distance measure one has to create a distance matrix using this distance matrix. Therefore, the next step is to create a distance matrix using Gower distance.

```{r gower}
# Create the distance matrix
source(paste(getwd(), "/R/02_gower-distance.R", sep = ""))

# Show a summary of the distance matrix
summary(gower_dist)
```

## Number of clusters

To determine the number of clusters a cluster analysis using PAM has been done with k ranging from 2 till 10. For each k the silhouette coefficient [@Rousseeuw1987] is calculated and the results plotted in the so called silhouette plot.

```{r silhouette-plot}
# Run the multiple cluster analysis for the silhouette plot
source(paste(getwd(), "/R/03_silhouette-plot.R", sep = ""))

kable(sil_width_df)
```

The silhouette coefficient is highest at k = 3 and k = 5. The k = 5 clusters solution is more meaningful. Therefore, 5 clusters are extracted.

```{r set_n_clusters}
# Set the number of clusters
n_clusters <- 5
```

## Cluster analysis

Next the cluster analysis is run, the used method is Partitioning Around Medoids [@Kaufman1990].

```{r pam}
# Run the cluster analysis using PAM
source(paste(getwd(), "/R/04_pam.R", sep = ""))
```

## Aggregate clustering variables

For the following analysis and interpretation the order of the clusters is quite confusing. Therefore the order of the clusters is changed to a meaningful order.

```{r cluster_order}
# Run the cluster analysis using PAM
source(paste(getwd(), "/R/05_change_cluster_order.R", sep = ""))
```

In this step the data that went into the cluster analysis is aggregated by cluster and shown in a table:

```{r results_aggregate}
# Aggregate the data for each cluster
source(paste(getwd(), "/R/06_aggregate.R", sep = ""))

# Show the data for each cluster in a table
kable(cluster_results, row.names = FALSE)
```

## Test for significant differences on each cluster variable

This step tests for a significant difference on each cluster variable. For continious variables the Kruskal-Wallis Rank Sum Test [@Kruskal1952] is used and for categorical variables Pearson's Chi-squared Test for Count Data [@Pearson1900] with Monte Carlo simulation [@Hope1968] is used. The number of repetitions used in the Monte Carlo simulation is 8000 [@Mundform2011].

```{r cluster_variables_significance}
# Test for significant difference on each cluster variable
source(paste(getwd(), "/R/07_tests_for_difference_in_cluster_variables.R", sep = ""))

# Show the test result for each variable in a table
kable(difference_tests_summary_cluster_variables)
```

For the variables where the difference between the clusters is significant a post-hoc test is conducted. For continious variables the post-hoc test used is the Conover-Iman test [@Conover1979] and for the categorical variables the residuals from the Pearson's Chi-squared Test for Count Data with Monte Carlo simulation is used [@Beasley1995]. Based on the residuals a p value is calculated. For this proces a self written R package called chisq.posthoc.test is used. In both cases the p values are adjusted using the Bonferroni method [@Dunn1961].

**Conover-Iman test for: `r cluster_results$Variable[1]`**
```{r difference_test_Q5_1}
# Conover-Iman test to determine which of the clusters differ.
conover.test(cluster_data$Q5_1,
             cluster_data$cluster,
             method = "bonferroni",
             altp = TRUE)
```

**Pearson's Chi-squared Test for Count Data with Monte Carlo simulation residuals analysis for: `r cluster_results$Variable[2]`**
```{r difference_test_Q5_2}
# Run a posthoc test based on the residuals to determine which clusters differ
kable(chisq.posthoc.test(
  table(cluster_data$Q5_2, cluster_data$cluster),
  method = "bonferroni",
  simulate.p.value = TRUE,
  B = 8000
))
```

**Pearson's Chi-squared Test for Count Data with Monte Carlo simulation residuals analysis for: `r cluster_results$Variable[3]`**
```{r difference_test_Q5_4}
# Run a posthoc test based on the residuals to determine which clusters differ
kable(chisq.posthoc.test(
  table(cluster_data$Q5_4, cluster_data$cluster),
  method = "bonferroni",
  simulate.p.value = TRUE,
  B = 8000
))
```

**Conover-Iman test for: `r cluster_results$Variable[4]`**
```{r difference_test_Q5_6}
# Conover-Iman test to determine which of the clusters differ.
conover.test(cluster_data$Q5_6,
             cluster_data$cluster,
             method = "bonferroni",
             altp = TRUE)
```

**Conover-Iman test for: `r cluster_results$Variable[5]`**
```{r difference_test_Q6_3}
# Conover-Iman test to determine which of the clusters differ.
conover.test(cluster_data$Q6_3,
             cluster_data$cluster,
             method = "bonferroni",
             altp = TRUE)
```

## Semester distribution

In this step the distribution of the students in the clusters over the semesters in which the data was gathered is shown.
```{r semester_count}
# Count the number of students by semester and by cluster
source(paste(getwd(), "/R/08_semester_count.R", sep = ""))

# Show the results in a table
kable(semester_count, row.names = FALSE)
```

Test if any semester is significantly more present in one of the semesters during which the data was gathered using Pearson's Chi-squared Test for Count Data.

```{r semester_count_sig_distribution}
# Pearson's Chi-squared test for the distribution of the semesters over the clusters
chisq.test(subset_df$Semester, subset_df$cluster)
```

## Aggregate dependent variables

In this step some variables from the original data set are added and aggregated by cluster to aid the interpretation.

```{r dependent}
# Aggregate the additional variables for the interpretation
source(paste(getwd(), "/R/09_filter.R", sep = ""))

# Show the data for each cluster in a table
kable(dependent_df, row.names = FALSE)
```

# Test for significant difference in dependent variables

This step tests for a significant difference on each dependent variable. For continious variables the Kruskal-Wallis Rank Sum Test is used and for categorical variables Pearson's Chi-squared Test for Count Data with Monte Carlo simulation is used.

```{r dependent_variables_significance}
# Test for significant difference on each dependent variable
source(paste(getwd(), "/R/10_tests_for_difference_in_dependent_variables.R", sep = ""))

# Show the test result for each variable in a table
kable(difference_tests_summary_dependent_variables)
```

For the variables where the difference between the clusters is significant a post-hoc test is conducted. For continious variables the post-hoc test used is the Conover-Iman test and for the categorical variables the residuals from the Pearson's Chi-squared Test for Count Data with Monte Carlo simulation is used. Based on the residuals a p value is calculated. For this proces a self written R package called chisq.posthoc.test is used. In both cases the p values are adjusted using the Bonferroni method.

**Conover-Iman test for: `r dependent_df$Variable[2]`**
```{r difference_test_Q2_2}
# Conover-Iman test to determine which of the clusters differ.
conover.test(
  dependent_df_raw$Q2_2,
  dependent_df_raw$cluster,
  method = "bonferroni",
  altp = TRUE
)
```

**Pearson's Chi-squared Test for Count Data with Monte Carlo simulation residuals analysis for: `r dependent_df$Variable[3]`**
```{r difference_test_Q3_1}
# Run a posthoc test based on the residuals to determine which clusters differ
kable(chisq.posthoc.test(
  table(dependent_df_raw$Q3_1, dependent_df_raw$cluster),
  method = "bonferroni",
  simulate.p.value = TRUE,
  B = 8000
))
```

**Conover-Iman test for: `r dependent_df$Variable[5]`**
```{r difference_test_Q8_1}
# Conover-Iman test to determine which of the clusters differ.
conover.test(
  as.numeric(dependent_df_raw$Q8_1),
  dependent_df_raw$cluster,
  method = "bonferroni",
  altp = TRUE
)
```

**Compare the answers about the learning process by cluster**
```{r lernprozess}
# Get the answers about the learning proces per cluster
source(paste(getwd(), "/R/11_proces.R", sep = ""))

# Show the data about the learning process for each cluster in a table
kable(learn_df, row.names = FALSE)
```

**Test significance for learning proces variables**

```{r learning_variables_significance}
# Test for significant difference on each dependent variable
source(paste(getwd(), "/R/12_tests_for_difference_in_learning_variables.R", sep = ""))

# Show the test result for each variable in a table
kable(difference_tests_summary_learning_variables)
```

**Pearson's Chi-squared Test for Count Data with Monte Carlo simulation residuals analysis for: `r learn_df$Variable[1]`**
```{r difference_test_Q6_1_1}
# Run a posthoc test based on the residuals to determine which clusters differ
kable(chisq.posthoc.test(
  table(learn_df_raw$Q6_1_1, learn_df_raw$cluster),
  method = "bonferroni",
  simulate.p.value = TRUE,
  B = 8000
))
```

**Pearson's Chi-squared Test for Count Data with Monte Carlo simulation residuals analysis for: `r learn_df$Variable[2]`**
```{r difference_test_Q6_1_2}
# Run a posthoc test based on the residuals to determine which clusters differ
kable(chisq.posthoc.test(
  table(learn_df_raw$Q6_1_2, learn_df_raw$cluster),
  method = "bonferroni",
  simulate.p.value = TRUE,
  B = 8000
))
```

**Pearson's Chi-squared Test for Count Data with Monte Carlo simulation residuals analysis for: `r learn_df$Variable[3]`**
```{r difference_test_Q6_1_3}
# Run a posthoc test based on the residuals to determine which clusters differ
kable(chisq.posthoc.test(
  table(learn_df_raw$Q6_1_3, learn_df_raw$cluster),
  method = "bonferroni",
  simulate.p.value = TRUE,
  B = 8000
))
```

**Pearson's Chi-squared Test for Count Data with Monte Carlo simulation residuals analysis for: `r learn_df$Variable[5]`**
```{r difference_test_Q6_1_5}
# Run a posthoc test based on the residuals to determine which clusters differ
kable(chisq.posthoc.test(
  table(learn_df_raw$Q6_1_5, learn_df_raw$cluster),
  method = "bonferroni",
  simulate.p.value = TRUE,
  B = 8000
))
```

# Session Info

Show the session info to state which packages have been used for the analysis.

```{r session}
# Show the session info
sessionInfo()
```

# References
